### 2021-05-14 Modules/sec_prlm_FIR_IIR.tex
After Israel's comments
Acronyms added

As stated in \cite{oppenheim1975digital} (and numerous other literature), two key filter types are Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filters.
This section will briefly overview the design methods and the analogy of ULA based beamforming to the temporal FIR filtering, as implied in Ch.~\ref{chap:intro}.

@@

    IIR filters may be viewed as a generalization of the FIR delay-and-sum architecture, where delayed instances of the array's output are fed back to the summation as in Fig.~\ref{fig_IIR_arch}, theoretically generating an infinite loop where each output sample is affected by all past input samples. 

@@

        tics=10,trim=0 0 0 0]{./Media/BASIC_FIR_FILTER_ARCH.png}
            \put (13, 36){\footnotesize{$s$}}
            \put (22, 22){\footnotesize{$\beta_{0}$}}
            \put (38.75, 22){\footnotesize{$\beta_{1}$}}
            \put (55.5, 22){\footnotesize{$\beta_{2}$}}
            \put (80.75, 22){\footnotesize{$\beta_{N-1}$}}
            \put (85, 10){\footnotesize{$z$}}

@@

        tics=10,trim=0 0 0 0]{./Media/BASIC_IIR_FILTER_ARCH.png}
            \put (60, 50){\footnotesize{$\beta_{0}$}}
            \put (60, 30){\footnotesize{$\beta_{1}$}}
            \put (60, 10){\footnotesize{$\beta_{2}$}}
            \put (36, 30){\footnotesize{$\alpha_{1}$}}
            \put (36, 10){\footnotesize{$\alpha_{2}$}}
            \put (10, 50){\footnotesize{$s$}}
            \put (85, 50){\footnotesize{$z$}}
            \put (47.5, 34.5){\footnotesize{$z^{-1}$}}
            \put (47.5, 14.5){\footnotesize{$z^{-1}$}}

@@

\par In the following, we revisit the basic interpretation of the ULA as the spatial equivalent to the temporal FIR filter \cite{van1988beamforming} where DOA is shown to be the matching spatial entity to the temporal frequency, hence the \emph{conventional beamformer} \cite{van2004optimum} is merely the spatial version of temporal domain FIR filtering. To this end, a brief overview on ULA beamforming and its resemblance to FIR filter design are presented next.
\par Consider an $N$-element ULA with inter-element spacing $d$, where its $n$'th sensor is positioned at $\vpi{n},\;\text{for}\; n=0,\ldots,N-1$. We set $\vpi{0}$ as the axis reference point and assume that the target of interest is positioned at $\vpt$.
Focusing on a far field localization problem, DOA and target range are to be estimated. Without loss of generality, aiming to simplify the exposition, we assume $2\text{D}$ planar problem, where DOA is described by a single angle $\thetaD$, measured from the array's broadside.

@@

Inspired by radar based applications, we also place a transmitter at $\vpi{0}$, assuming the transmitted signal $s$, is reflected back from the target and re-impinges the array, with a total time delay of $\tau_{\text{pd}}=2R/c$ seconds, where $R = \norm{\vpt-\vpi{0}}$ is the target's range and $c$ represents the propagation velocity of the signal in the medium.
\par Let $x_{n}(t)$ be the measured signal at the $n$'th sensor

@@

x_{n}(t) = gs\Brack{t-\tau_{pd}-\tau_{n}},

@@

where $\tau_{n}=nd\cos\Brack{\thetaD}/c$ represents the time difference of arrival between the $n$'th sensor and the reference sensor and $g$, being a scalar in an anechoic environment, is the channel's gain, related to both propagation and the target's radar cross section (RCS).

@@

\vecnot{\F{x}}\rBrace{\omega,\theta_{d}}=g\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}\rBrace{\theta}}}

@@

    Z\rBrace{\omega,\theta_{d}} = g\vecnot{\beta}^{H}\rBrace{\omega}\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}.

@@

\theta=\omega{d\cos\Brack{\thetaD}}/{c},

@@

Z\rBrace{\omega,\theta} = g\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}\sum_{n=0}^{N-1}\beta^{*}_{n}\rBrace{\omega}\exp\Brack{-jn\theta},

@@

hence in the ULA case, aiming for a desired spatial response, the weights vector $\vecnot{\beta}\rBrace{\omega}$ configuration is mathematically equivalent to an FIR filter design~\cite{van1988beamforming,benesty2018}. 
Assuming narrowband stimuli signals, we suppress $\omega$ dependency in the notation throughout the rest of this paper, where possible.
\par It is well known \cite{rabiner1974some} that IIR filters have several appealing advantages over their matching (in terms of number of taps) FIR counterparts.
The main advantage of IIR filters is that for a given set of frequency response specifications, the IIR filter will (in most cases) require substantially less coefficients, which may even reach few order of magnitude \cite{rabiner1974some}.

@@

However, a well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant, namely linearly dependent in the phase response derivative of the specific frequency.
Therefore, linear phase implies that the phase response's derivative (i.e. signal latency) is constant which implies all input frequencies will reside in the processor for the exact same time, preventing dispersion which practically causes pulse widening.
Due to this important property, many applications which are sensitive to signal distortions, especially communication based systems, chose the FIR architecture, for it is possible to design it as linear phase filter.
Acquainted with the analogy between frequency and DOA, it seems that when localization applications are considered, where the DOA resolution criteria is more important than the consistency of detection latency between separate DOAs, the linear phase property is less relevant.

@@

\par In standard radar signal processing schemes, a waveform is transmitted to, and reflected from the target of interest. Then, the reflected signal is processed by the radar reception array in order to estimate the target's dynamics (e.g., DOA, range, velocity etc.).
As opposed to the standard scheme, we suggest in Chapter.~\ref{chap:firstchap} a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.
Another deviation from traditional radar processing, used to simplify the exposition, is using continuous-wave (CW) stimuli, rather than using pulse based signals.

>>>>

As stated in \cite{oppenheim1975digital} (and numerous other sources), two key filter types are FIR and IIR filters.
This section will briefly overview the design methods and the analogy of ULA based beamforming to the temporal FIR filtering, as implied in Chapter.~\ref{chap:intro}.

@@

    IIR filters may be viewed as a generalization of the FIR delay-and-sum architecture, where delayed instances of the array's output are fed back to the summation as in Fig.~\ref{fig_IIR_arch}, generating an infinite loop where each output sample is affected by all past input samples. 

@@

        tics=10,trim=0 0 0 0]{./Media/FIR_arch.png}
            \put (13, 38){$s$}
            \put (26, 33.5){$Z^{-1}$}
            \put (47, 33.5){$Z^{-1}$}
            \put (68, 33.5){$Z^{-1}$}
            \put (37.5, 3){$+$}
            \put (59, 3){$+$}
            \put (80, 3){$+$}
            \put (20.5, 18){$\beta_{0}$}
            \put (42, 18){$\beta_{1}$}
            \put (63, 18){$\beta_{2}$}
            \put (84, 18){$\beta_{N-1}$}
            \put (90, 7){$z$}

@@

        tics=10,trim=0 0 0 0]{./Media/IIR_arch.png}
            \put (60, 55){$\beta_{0}$}
            \put (60, 30){$\beta_{1}$}
            \put (60, 6){$\beta_{2}$}
            \put (39, 30){$\alpha_{1}$}
            \put (39, 6){$\alpha_{2}$}
            \put (10, 54){$s$}
            \put (90, 54){$z$}
            \put (48.5, 37.5){$z^{-1}$}
            \put (48.5, 13){$z^{-1}$}
            \put (21, 25){+}
            \put (21, 50){+}
            \put (78, 25){+}
            \put (78, 50){+}

@@

\par In the following, we revisit the basic interpretation of the ULA as the spatial equivalent to the temporal FIR filter \cite{van1988beamforming} where DOA is shown to be the matching spatial entity to the temporal frequency, hence the CB \cite{van2004optimum} is merely the spatial version of temporal domain FIR filtering. 
% To this end, a brief overview on ULA beamforming and its resemblance to FIR filter design are presented next.
\par 
% Consider an $N$-element ULA with inter-element spacing $d$, where its $n$'th sensor is positioned at $\vpi{n},\;\text{for}\; n=0,\ldots,N-1$. We set $\vpi{0}$ as the axis reference point and assume that the target of interest is positioned at $\vpt$.
We consider the same ULA as in Sec.~\ref{sec:prlm_propWaveField}, where a target of interest is placed at $\vpt$.
Focusing on a far field localization problem, DOA and target range are to be estimated. 

@@

% In the following, we place a transmitter is at $\vpi{0}$, which emits the signal $s$.
% Also, we also consider a reflective target, with distance $R = \norm{\vpt-\vpi{0}}$ from the array.
% Then, the reflected signal re-impinges the array, with a total time delay of $\tau_{\text{pd}}=2R/c$ seconds, where $c$ represents the propagation velocity of the signal in the medium.
\par As in RADAR applications, the target is assumed to emit a signal $s\rBrace{t}$, where for convenience $s\rBrace{t=0}$ is the signal which impinges $\vpi{0}$ .
Let $x_{n}(t)$ be the measured signal at the $n$'th sensor

@@

x_{n}(t) = s\Brack{t-\tau_{n}},

@@

where, $\tau_{n}=n\tau_{\theta_{d}}=nd\cos\Brack{\thetaD}/c$ represents the time difference of arrival between the $n$'th sensor and the reference sensor.
% and $g$, being a scalar in an anechoic environment, is the channel's gain, related to both propagation and the target's radar cross section (RCS).

@@

\vecnot{\F{x}}\rBrace{\omega,\theta_{d}}=\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}

@@

    Z\rBrace{\omega,\theta_{d}} = \vecnot{\beta}^{T}\rBrace{\omega}\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}.

@@

\theta=\omega\tau_{\theta_{d}},

@@

Z\rBrace{\omega,\theta} = \F{s}\rBrace{\omega}\sum_{n=0}^{N-1}\beta_{n}\rBrace{\omega}\exp\Brack{-jn\theta},

@@

hence in the ULA case, aiming for a desired spatial response, the weights vector $\vecnot{\beta}\rBrace{\omega}$ configuration is mathematically equivalent to an FIR filter design~\cite{van1988beamforming,benesty2018} as illustrated in Fig.~\ref{fig_analogyULAFIR}.
        tics=10,trim=0 0 0 0]{./Media/analogyFIRULA.png}
            \put (19.5, 23){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (42.25, 23){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (55.25, 23){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (70, 23){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (91, 70){$s\rBrace{t}$}
            \put (85, 72.5){$\vpi{t}$}
    \caption{An illustration of the analogy between ULA based spatial processing and an FIR temporal filter. The propagation of a plane wave from one sensor to its neighbour (solid tilted lines) is equivalent to the temporal time shift in temporal digital processors (dashed lines and rectangular taps).}
    \label{fig_analogyULAFIR}
Assuming narrowband stimuli signals, we suppress $\omega$ dependency in the notation throughout the rest of this thesis, where possible.
\par It is well known \cite{rabiner1974some} that IIR filters have several appealing advantages over their matching FIR counterparts.
The main advantage of IIR filters over their FIR counterparts, is their efficiency in terms of the required sensors quantity, which even reaches few orders of magnitude \cite{rabiner1974some} in some cases.

@@

However, a well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant.
When the dependence is linear, it implies that the phase response derivative (i.e. signal latency) is constant such that there is no dispersion.
Due to this important property, applications which are sensitive to signal distortions, especially communication based systems, are based on FIR architecture, for it is possible to design it as linear phase filter.
Also notable is the stability issues of the IIR filter.
% Acquainted with the analogy between frequency and DOA, it seems that when localization applications are considered, where the DOA resolution criteria is more important than the consistency of detection latency between separate DOAs, the linear phase property is less relevant.
% This observation, together with the IIR's efficiency motivate the search for the spatial structure which will be analogous to the IIR filter.
Motivated by the IIR efficiency, we wish to find the spatial structure that will be analogous to the IIR filer design.
To this end, we aim to incorporate a spatial feedback which will serve as the temporal feedback in the IIR architecture ($\alpha_{1}, \alpha_{2}$ in Fig.~\ref{fig_IIR_arch}) as illustrated in Fig.~\ref{fig_analogyIIRFB}.
        tics=10,trim=0 0 0 0]{./Media/analogyIIRFB.png}
            \put (17.5, 20.5){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (38.25, 20.5){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (50, 20.5){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (63.25, 20.5){\footnotesize{$Z^{-\tau_{\theta_{d}}}$}}
            \put (94.5, 63.5){$s\rBrace{t}$}
            \put (78, 65.5){$\vpi{t}$}
    \caption{An illustration of the spatial feedback concept. On top of the FIR equivalence (non-bold dashed lines), the incorporation of the feedback is presented with the bold dashed lines, generating the ``IIR`` part of the spatial processor. The multipliers with the bold rectangle wrappers act as the feedback coefficients ($\vecnot{\alpha}$) of Fig.~\ref{fig_IIR_arch}.}
    \label{fig_analogyIIRFB}

@@

\par In standard radar signal processing schemes, a waveform is transmitted to and reflected from the target of interest. 
Then, the reflected signal is processed by the radar reception array in order to estimate the target's dynamics (e.g., DOA, range, velocity etc.).
As opposed to the standard scheme (as implied from Fig.~\ref{fig_analogyIIRFB}) we suggest a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.
% Another deviation from modern radar processing, used to simplify the exposition, is using continuous-wave (CW) stimuli, rather than using pulse based signals.

### 2021-05-14 Modules/sec_prlm_array_perf.tex
Acronyms added

Those criteria are then used to quantify the presented architecture's performance in Chapter.\ref{chap:firstchap}.

@@

Considering DOA estimators, the main tool for assessing an array's localization performance is expressing it's response to impinging signals with respect to the signal's DOA.

@@

  In both plots, a comparison between the conventional beamformer's responses of 3,6 and 12 elements arrays is emulated. 

@@

Although for sensors in 3D space the DOA may depend on two angles (azimuth and elevation), for the sake of simplicity, we consider a planar (i.e. azimuth without elevation) DOA-only dependant array response (i.e. the \emph{beampattern})

@@

The analysis can be done theoretically, when the beampattern's mathematical expression is known or it can be conducted numerically for empirical scenarios.

@@

The DOAs of the impinging signal, may be presented in various ways where units may be degrees or radians.

@@

Also, we choose to plot beampatterns using $\Delta\theta = \theta\rBrace{\theta_{d}} - \theta\rBrace{\theta_{s}}$, where $\theta_{s}$ is the steering angle of the array, for it centers the mainlobe, enabling convenient comparison.

@@

This quantity will be referred to as $\mathcal{H}$.

@@

For standard ULA, assuming large $N$ values, it is known \cite{van2004optimum} that

@@

Obviously, we aim to show that the presented architecture increases the resolution, comparing to other arrays by proving that the beamwidth shrinks.

@@

A known \cite{van2004optimum} parasitic phenomenon of beampatterns is sidelobes, which manifests as energy peaks outside the mainlobe, typically highest near the mainlobe and decreasing towards the deges of the beampattern.

@@

Related spatial performance metrics are the sidelobes gain and rate of decrease

@@

As in \cite{van2004optimum}, in a transmitting array or aperture, $\mathcal{D}$ represents the maximum radiation intensity (power per DOA) divided by the average radiation intensity (averaged over all DOAs).
Also, in a receiving antenna, the denominator represents the noise power at the array (or aperture) output due to isotropic noise (noise distributed uniformly over all DOAs). 
The numerator will represent the power due to a signal arriving from a certain DOA ($\theta_{T}$) - Thus, $\mathcal{D}$ can be interpreted as the array gain against isotropic noise.

>>>>

Those criteria are then used to quantify the presented architecture's performance in Chapter.~\ref{chap:firstchap}.

@@

Considering DOA estimators, the main tool for assessing an array's localization performance is expressing it's response to impinging signals with various DOA.

@@

  In both plots, a comparison between the CB's responses of 3,6 and 12 elements arrays is simulated. 

@@

For sensors in 3D space the DOA depends on two angles (azimuth and elevation).
However, for the sake of simplicity, we consider a planar (i.e. azimuth without elevation) DOA such that the beampattern is a function of an angle $\theta$

@@

The analysis can be done theoretically, when the beampattern's mathematical expression is known or it can be conducted numerically.

@@

The DOAs of the impinging signal may be presented in various ways where units may be degrees or radians.

@@

Also, we choose to plot beampatterns using $\Delta\theta = \theta\rBrace{\theta_{d}} - \theta\rBrace{\theta_{s}}$, where $\theta_{s}$ is the steering angle of the array, it centers the mainlobe, enabling convenient comparison.

@@

This quantity will be referred to as $$\mathcal{H}\rBrace{\theta} = H\rBrace{\theta}/H\rBrace{\theta_{s}}$$ where $H$ is the array's response and $\theta_{s}$ is the DOA that the array is steered to.

@@

For standard $N$-element ULA, assuming large $N$ values, it is known \cite{van2004optimum} that

@@

Obviously, we aim to show that the presented architecture increases the resolution, compared to other arrays by proving that the beamwidth shrinks below the $1.4/N$ limit.

@@

A known \cite{van2004optimum} parasitic phenomenon of beampatterns is sidelobes, which manifests as energy peaks outside the mainlobe, typically highest near the mainlobe and decreasing towards the edges of the beampattern.

@@

Related spatial performance metrics are the sidelobes gain and rate of decrease in sidelobe gain.

@@

As in \cite{van2004optimum}, in a transmitting array, $\mathcal{D}$ represents the maximum radiation intensity (power per DOA) divided by the average radiation intensity (averaged over all DOAs).
Also, in a receiving array, the denominator represents the noise power at the array output due to isotropic noise (noise distributed uniformly over all DOAs). 
The numerator will represent the power due to a signal arriving from a certain DOA ($\theta_{d}$) - Thus, $\mathcal{D}$ can be interpreted as the array gain against isotropic noise.

### 2021-05-14 Modules/sec_prlm_localization.tex
Acronyms added

In the wide and very active research field of parametric estimation, one subject of very high relevance to this thesis is the sensor-array based localization - i.e. positional parameters estimation of impinging signal which is measured by an array of sensors.

@@

limitations, namely ``its performance, is directly dependent upon the physical size of the array (the aperture), regardless of the available data collection time and signal-to-noise ratio (SNR)`` \cite{krim1996two}.

@@

In this contribution we tackle this issue as will shown in Chapter.~\ref{chap:firstchap} by introducing a spatial feedback.

@@

The later approach, of parameter estimation assumes a physical model of the experimental scenario, which serve a basis for the development of spatio-temporal estimators which take advantage not only from the spatial diversity but also from the temporal information collected via sequential time separated snapshots in order to increase the overall SNR.

>>>>

In the wide and very active research field of parametric estimation, one issue of relevance to this thesis is the sensor-array based localization - i.e. positional parameters estimation of impinging signal which is measured by an array of sensors.

@@

limitations, namely ``its performance, is directly dependent upon the physical size of the array (the aperture), regardless of the available data collection time and signal-to-noise ratio`` \cite{krim1996two}.

@@

In this contribution we tackle this issue as will be shown in Chapter.~\ref{chap:firstchap} by introducing a spatial feedback.

@@

The later approach, of parameter estimation assumes a physical model of the experimental scenario.
Then, spatio-temporal estimators take advantage of not only the spatial diversity but also from the temporal information in order to increase the overall SNR.

### 2021-05-14 Modules/sec_prlm_propWaveField.tex
Acronyms added

where $ \vecnot{x} = \vBrace{x,y,z}^{T} $ emphasizes the fact that for a given time, $t_{0}$, all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$ are with the same phase value where $\vecnot{k}\vecnot{x}=constant$ are planes of constant phase value.

@@

The spatial diversity of the ULA elements is expressed as a TOA difference of the impinging signal in each sensor which will be shown to provide clues for the signal's DOA as can be seen in Fig.~\ref{fig_ULA_sketch}.

@@

        \put(33.25,2.25){\rotatebox{0}{\tiny{$d$}}}
        \put(16,26){\rotatebox{-36}{\tiny{$c\tau=d\cos{\theta}$}}}
        \put(45.5,15.75){\rotatebox{0}{$\theta$}}
        \put(15,7){\rotatebox{0}{\tiny{0}}}
        \put(28,7){\rotatebox{0}{\tiny{1}}}
        \put(41,7){\rotatebox{0}{\tiny{2}}}
        \put(80,7){\rotatebox{0}{\tiny{$N-1$}}}
        \put(2,24){\rotatebox{-36}{\tiny{$t=0$}}}
        \put(2,34){\rotatebox{-36}{\tiny{$t=\tau$}}}
        \put(2,44){\rotatebox{-36}{\tiny{$t=2\tau$}}}
        \put(40,44){\rotatebox{-36}{\tiny{$t=\rBrace{N-1}\tau$}}}

>>>>

where $ \vecnot{x} = \vBrace{x,y,z}^{T} $ emphasizes the fact that for a given time, $t_{0}$, all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$ are with the same phase value where $\vecnot{k}^{T}\vecnot{x}=constant$ are planes of constant phase value.

@@

The spatial diversity of the ULA elements is expressed as a TOA difference of the impinging signal in each sensor which will be shown to provide clues for the signal's DOA (measured from the array's broadside) as can be seen in Fig.~\ref{fig_ULA_sketch}.

@@

        \put(57.5,2.25){\rotatebox{0}{\tiny{$d$}}}
        \put(42,34){\rotatebox{-36}{\tiny{$c\tau=d\cos{\theta}$}}}
        \put(19,15.75){\rotatebox{0}{$\theta$}}
        \put(15,7){\rotatebox{0}{\tiny{$N-1$}}}
        \put(65,7){\rotatebox{0}{\tiny{1}}}
        \put(52,7){\rotatebox{0}{\tiny{2}}}
        \put(80,7){\rotatebox{0}{\tiny{0}}}
        \put(-5,29){\rotatebox{-36}{\tiny{$t=-\rBrace{N-1}\tau$}}}
        \put(13,42){\rotatebox{-36}{\tiny{$t=-2\tau$}}}
        \put(27,42){\rotatebox{-36}{\tiny{$t=-\tau$}}}
        \put(42,42){\rotatebox{-36}{\tiny{$t=0$}}}

### 2021-05-14 Modules/sec_prlm_sensorArrays.tex
Acronyms added

Sensor arrays are sets of independently positioned sensors, where each sensor is sampling temporal snapshots of impinging signals.
The samples from the entire array are then fused together in order to extract the underlying data.
The spatial diversity of the sampled data allows the extraction of spatial features.
Sensor arrays are used in numerous applications, ranging from source localization, communication, medical applications, astronomy etc.
\par 
When designing a sensor array, multiple consideration are to be taken into account, ranging from physical characteristics e.g. area, weight and carrying platform, through performance related considerations such as accuracy, spatial selectivity, signal to noise ratio and even the financial aspect.
The most fundamental array specification is its geometry, for it dictates the spatial relations between simultaneous samples measured by the array's sensors, which also greatly influences the processing methods of the raw-data.
A basic example of a sensor array is the ULA (see Fig.~\ref{fig_ULA}), where its elements are uniformly spaced on a straight line with distance $d$ between each pair of sensors.
Considering an array of $N$ elements, the sensors are positioned at
\begin{equation}
p_{n}=nd,\ n=0,\dots,N-1
\end{equation}
where $n$ is the sensors index; $p_{n}$ denotes the position of the $n$'th sensor, $p_{0} = 0$ corresponds to the left-hand side of the array as illustrated in Fig.~\ref{fig_ULA}.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayBasic.png}
        \put (4, 4.5) {\tiny{$N$}}
        \put (86, 4.5) {\tiny{$N-1$}}
        \put (83.25, 28.5) {\tiny{$\rBrace{N-1}\cdot{}d$}}
        \end{overpic}
    \end{center}
     \caption{A uniform linear array of size $N$ and spacing $d$.}
    \label{fig_ULA}
\end{figure}
\par To properly formulate the discussed additional spatial information when using sensor arrays, a quick overview on wave propagation is due and presented in the following section (Sec.~\ref{sec:prlm_propWaveField}).

>>>>

Sensor arrays are sets of independently positioned sensors, where each sensor is sampling temporal snapshots of impinging signals.
The samples from the entire array are then fused together in order to extract the underlying data.
The spatial diversity of the sampled data allows the extraction of spatial features.
Sensor arrays are used in numerous applications, ranging from source localization, communication, medical applications, astronomy etc.
\par 
When designing a sensor array, multiple consideration are to be taken into account, ranging from physical characteristics e.g. area, weight and carrying platform, through performance related considerations such as accuracy, spatial selectivity, SNR and even the financial aspect.
The most fundamental array specification is its geometry, for it dictates the spatial relations between simultaneous samples measured by the array's sensors, which also greatly influences the processing methods of the raw-data.
A basic example of a sensor array is the ULA (see Fig.~\ref{fig_ULA}), where its elements are uniformly spaced on a straight line with distance $d$ between each pair of sensors.
Considering an array of $N$ elements, the sensors are positioned at
\begin{equation}
p_{n}=nd,\ n=0,\dots,N-1
\end{equation}
where $n$ is the sensors index; $p_{n}$ denotes the position of the $n$'th sensor, $p_{0} = 0$ corresponds to the left-hand side of the array as illustrated in Fig.~\ref{fig_ULA}.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayBasic.png}
        \put (4, 4.5) {\tiny{$N$}}
        \put (86, 4.5) {\tiny{$N-1$}}
        \put (83.25, 28.5) {\tiny{$\rBrace{N-1}\cdot{}d$}}
        \end{overpic}
    \end{center}
     \caption{A ULA of size $N$ and spacing $d$.}
    \label{fig_ULA}
\end{figure}
\par To properly formulate the discussed additional spatial information when using sensor arrays, a quick overview on wave propagation is due and presented in the following section (Sec.~\ref{sec:prlm_propWaveField}).

### 2021-05-14 Modules/sec_prlm_timeSpaceSig.tex
Acronyms added

As mentioned before, the measurements of the sensor array's elements are fused together, where spatial information resides in the relations between samples taken the same time $t$.
A basic application of the sensor array is the beamformer, designed to enhance signals from certain directions while suppressing other signals from unwanted directions, forming a beam directed to the enhanced direction.  
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.6\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayProc.png}
        \put(45,33.25){\rotatebox{0}{$\theta$}}
        \put(15,25){\rotatebox{0}{\tiny{0}}}
        \put(28,25){\rotatebox{0}{\tiny{1}}}
        \put(41,25){\rotatebox{0}{\tiny{2}}}
        \put(80,25){\rotatebox{0}{\tiny{$N-1$}}}
        \put(16,20){\rotatebox{0}{\tiny{$\omega_{0}$}}}
        \put(29.5,20){\rotatebox{0}{\tiny{$\omega_{1}$}}}
        \put(42,20){\rotatebox{0}{\tiny{$\omega_{2}$}}}
        \put(81,20){\rotatebox{0}{\tiny{$\omega_{N-1}$}}}
        \put(50.5,3.5){\rotatebox{0}{\tiny{$z\rBrace{t}$}}}
        \end{overpic}
    \end{center}
     \caption{ULA based coherent processing of simultaneous spatially sampled signal, measuring a wavefront impinging from DOA $\theta$}
    \label{fig_ULA_imping}
\end{figure}
In complimentary to Fig.~\ref{fig_ULA_sketch}, Fig.~\ref{fig_ULA_imping} adds the processor layer of the beamformer.
\par A temporal snapshot measured by the array at time $t$ is
\begin{equation}
\vecnot{f}\rBrace{t,\vecnot{p}} = \vBrace{f\rBrace{t,p_{0}},\dots,f\rBrace{t,p_{N-1}}}^{T},
\end{equation}
where $p_{n}$ denotes the $n$-th sensor position, $t$ denotes the continuous-time variable and $^{T}$ denotes the transpose operation. 
Considering the \emph{conventional beamformer} \cite{van2004optimum}, the output ($z\rBrace{t}$) is expressed by:
\begin{equation}
z\rBrace{t} = \sum_{n=0}^{N-1}{f\rBrace{t,p_{n}}\omega_{n}^{*}},
\end{equation}
where $\omega_{n}$ is a complex weight used for the $n$'th sensor's output, as shown in Fig.~\ref{fig_ULA_imping} and $^{*}$ denotes complex conjugation. 
Consider $s\rBrace{t} = e^{j\omega{t}}$ as a plane wave propagating at temporal frequency $\omega$, and $\theta \in \vBrace{-\frac{\pi}{2},\frac{\pi}{2}}$ as the direction of arrival (DOA) angle measured with respect to the broadside of the linear array, as shown in Fig.~\ref{fig_ULA_imping}.
The wave signal, spatially sampled by the sensor array inputs is
\begin{equation}
\label{eq_prlm_timeSpace_outputVec}
\vecnot{f}\rBrace{t,\vecnot{p}} = \vBrace{s\rBrace{t},s\rBrace{t-\tau}\dots, s\rBrace{t-\rBrace{N-1}\tau}}^{T},
\end{equation}
where $\tau = \frac{d\cdot{}\cos{\theta}}{c}$ is the propagation delay between consecutive sensors and $c$ is the wave's velocity in the medium.
Considering narrowband signals, the time delay $\tau$ translates to a mere phase shift and the \emph{steering vector} takes the form of
\begin{equation}
\vecnot{d}\rBrace{\theta}=\vBrace{1,e^{-j\omega\tau},\dots,e^{-j\omega\tau\rBrace{N-1}}}^{T}.
\end{equation}
As every impinging signal will cause the excitation of the array elements in the form of a specific steering vector (up to a common additive phase), we define the \emph{array manifold} to be the set of all steering vectors of $\theta\in\left[0,2\pi\right)$, hence it also spans the received signal subspace in the noiseless scenario.

>>>>

As mentioned before, the measurements of the sensor array's elements are fused together, where spatial information resides in the relations between samples taken the same time $t$.
A basic application of the sensor array is the beamformer, designed to enhance signals from certain directions while suppressing other signals from unwanted directions, forming a beam directed to the enhanced direction.  
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.6\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayProc.png}
        \put(19.5,29.5){\rotatebox{0}{$\theta$}}
        \put(15,21.5){\rotatebox{0}{\tiny{$N-1$}}}
        \put(65,21.5){\rotatebox{0}{\tiny{1}}}
        \put(52,21.5){\rotatebox{0}{\tiny{2}}}
        \put(80,21.5){\rotatebox{0}{\tiny{0}}}
        \put(16,18){\rotatebox{0}{\tiny{$\omega_{N-1}$}}}
        \put(54,18){\rotatebox{0}{\tiny{$\omega_{2}$}}}
        \put(67,18){\rotatebox{0}{\tiny{$\omega_{1}$}}}
        \put(81,18){\rotatebox{0}{\tiny{$\omega_{0}$}}}
        \put(50.5,3.5){\rotatebox{0}{\tiny{$z\rBrace{t}$}}}
        \end{overpic}
    \end{center}
     \caption{ULA based coherent processing of simultaneous spatially sampled signal, measuring a wavefront impinging from DOA $\theta$}
    \label{fig_ULA_imping}
\end{figure}
In complimentary to Fig.~\ref{fig_ULA_sketch}, Fig.~\ref{fig_ULA_imping} adds the processor layer of the beamformer.
\par A temporal snapshot measured by the array at time $t$ is
\begin{equation}
\vecnot{f}\rBrace{t,\vecnot{p}} = \vBrace{f\rBrace{t,p_{0}},\dots,f\rBrace{t,p_{N-1}}}^{T}.
\end{equation}
Considering the CB \cite{van2004optimum}, the output ($z\rBrace{t}$) is expressed by:
\begin{equation}
z\rBrace{t} = \sum_{n=0}^{N-1}{f\rBrace{t,p_{n}}\omega_{n}^{*}},
\end{equation}
where $\omega_{n}$ is a complex weight used for the $n$'th sensor's output, as shown in Fig.~\ref{fig_ULA_imping} and $^{*}$ denotes complex conjugation. 
Consider $s\rBrace{t} = e^{j\omega{t}}$ as a plane wave propagating at temporal frequency $\omega$, and $\theta \in \vBrace{-\frac{\pi}{2},\frac{\pi}{2}}$ as the DOA angle measured with respect to the broadside of the linear array, as shown in Fig.~\ref{fig_ULA_imping}.
The wave signal, spatially sampled by the sensor array inputs is
\begin{equation}
\label{eq_prlm_timeSpace_outputVec}
\vecnot{f}\rBrace{t,\vecnot{p}} = \vBrace{s\rBrace{t},s\rBrace{t-\tau}\dots, s\rBrace{t-\rBrace{N-1}\tau}}^{T},
\end{equation}
where $\tau = \frac{d\cdot{}\cos{\theta}}{c}$ is the propagation delay between consecutive sensors and $c$ is the wave's velocity in the medium.
Considering narrowband signals, the time delay $\tau$ translates to a mere phase shift and the \emph{steering vector} takes the form of
\begin{equation}
\vecnot{d}\rBrace{\theta}=\vBrace{1,e^{-j\omega\tau},\dots,e^{-j\omega\tau\rBrace{N-1}}}^{T}.
\end{equation}
As every impinging signal will cause the excitation of the array elements in the form of a specific steering vector (up to a common additive phase), we define the \emph{array manifold} to be the set of all steering vectors of $\theta\in\left[0,2\pi\right)$, hence it also spans the received signal subspace in the noiseless scenario.
